---
title: "Homework06"
author: "Emma Estabrook"
date: "2/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Open Libraries
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

# Part 1

## Fake Data
```{r}
# quick and dirty, a truncated normal distribution to work on the solution set

#z <- rnorm(n=3000,mean=0.2)
#z <- data.frame(1:3000,z)
#names(z) <- list("ID","myVar")
#z <- z[z$myVar>0,]
#str(z)
#summary(z$myVar)
```

## Real Data
```{r}
z <- read.table("SurveyVolumeEstimations.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
str(z)
summary(z)
```

## Plot histogram of data
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

## Add empirical density curve
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75) # smooths out the shape of the histogram. It does not assume any particular probability distribution
print(p1)
```

## Get maximum likelihood parameters for normal
```{r}
# This fits a normal distribution to the data
# maximum likelihood estimator of the parameters of the normal (mean and variance)
normPars <- fitdistr(z$myVar,"normal") # this is how we get the maximm likelihood parameters for a normal distribution fitted to the data
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

## Plot normal probability density
```{r}
meanML <- normPars$estimate["mean"] #data from fitdistr
sdML <- normPars$estimate["sd"] #data from fitdistr


xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, 
                          y = ..y..), 
                      fun = dnorm,
                      colour="red", 
                      n = length(z$myVar),
                      args = list(mean = meanML, sd = sdML))
p1 + stat # plots to graph
```
 
## Plot exponential probability density
```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```

## Plot uniform probability density
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

## Plot gamma probability density
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```
 
## Plot beta probability density
```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

# Part 2
### Maximum liklihood parameters
```{r}
parameters <- fitdistr(z$myVar,"gamma")
print(parameters)

shapeML <- parameters$estimate["shape"]
rateML <- parameters$estimate["rate"]
```

### Simulated Data
```{r}
sim_myVar <- data.frame(ID=seq(1:11),myVar= rgamma(11,shape=shapeML,rate=rateML))

p2 <- ggplot(data=sim_myVar, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p2)

p2 <-  p2 +  geom_density(linetype="dotted",size=0.75)
print(p2)
```

### Real Data
```{r}
print(p1)
```

## Write Up
I think the model did a good job at simulating the real data which was surprising for me. Having only 11 collections with none of the curves lining up super well, I was expecting the simulation to fail. It is doing a good job because of the high density bar. 