---
title: "Homework 05"
author: "Emma Estabrook"
date: "2/12/2020"
output: html_document
---

## Working with matrices, lists, and data frames

1. Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
n_dims <- runif(1,min=3,max=10)
print(n_dims)
```

* Create a vector of consecutive integers from 1 to n_dims2.
```{r}
x <- seq(1,n_dims^2)
print(x)
```

* Use the sample function to randomly reshuffle these values.
```{r}
y <- sample(x)
print(y)
```

* create a square matrix with these elements.
```{r}
m <- matrix(data=y,nrow=n_dims,ncol=n_dims)
```

* print out the matrix.
```{r}
print(m)
```

* find a function in r to transpose the matrix.
```{r}
m2 <- t(m)
```

* print it out again and note how it has changed.
```{r}
print(m2)
```

* calculate the sum and the mean of the elements in the first row and the last row.
```{r}
row1 <- c(sum(m2[1,]),
          mean(m2[1,]))
print(row1) # sum and average of row 1

row4 <- c(sum(m2[4,]),
          mean(m2[4,]))
print(row4) # sum and average of row 4
```

* read about the eigen() function and use it on your matrix
```{r}
m3 <- eigen(m2, only.values = FALSE, EISPACK = FALSE)
print(m3)
```

* look carefully at the elements of $values and $vectors. What kind of numbers are these?
```
$vectors and $values are positive and negative numbers. 
```

* dig in with the typeof() function to figure out their type.
```{r}
typeof(m3)
```

* if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.

2. Create a list with the following named elements:
* my_matrix, which is a 4 x 4 matrix filled with random uniform values
* my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
* my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
# my_matrix
my_matrix <- matrix(runif(16), nrow=4, ncol=4)
print(my_matrix)

# my_logical
a <- runif(100)
print(a)
my_logical <- a > 0.2
print(my_logical)

# my_letters
a2 <- letters[1:26]
print(a2)

my_letters <- sample(a2, 26)
print(my_letters)

# List
list1 <- list(my_matrix, my_logical, my_letters)
print(list1)

```

* create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
my_matrix2 <- matrix(my_matrix[2,2])
print(my_matrix2)

my_logical2 <- c(my_logical[2])
print(my_logical2)

my_letters2 <- c(my_letters[2])
print(my_letters2)

list2 <- list(my_matrix2, my_logical2, my_letters2)
print(list2)
```

*use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(list2[[1]])
typeof(list2[[2]])
typeof(list2[[3]])

```

* combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
vector <- c(list2[[1]], list2[[2]], list2[[3]])
print(vector)
```

what is the data type of this vector?
```{r}
typeof(vector)
```


3. Create a data frame with two variables (= columns) and 26 cases (= rows).
* call the first variable my_unis and fill it with a random uniform value from 0 to 10
* call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
my_unis <- runif(26, min=0, max=10)

a3 <- LETTERS[1:26]
my_letters <- sample(a3, 26)

dFrame <- data.frame(my_unis, my_letters)
print(dFrame)
```
* for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
```{r}
dFrame[sample(my_unis,4, replace = FALSE),1] <- NA
print(dFrame)
```

* for the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
is.na.data.frame(dFrame)
```

* for the second variable, sort it in alphabetical order
```{r}
dFrame[order(dFrame$my_letters), c(1,2)]
```


* calculate the column mean for the first variable.
```{r}
mean <- mean(my_unis)
print(mean)
```


